---
title: "Dijkstras Practice"
output: html_document
date: "2023-02-22"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(gtools)
```

# Creating Example Problem

##Step 1: Create the data frame of NODES
```{r}
nodes <- data.frame(
  id = c("95758", "97752", "93405", "90210"),
  node_type = c("home", "airport", "bus", "office")
)
head(nodes)
```

##Step 2: Create data frame of EDGES
```{r}
# Sample data frame
df <- data.frame(
  from = c("95758", "95758", "93405", "97752", "95758"),
  to = c("90210", "93405", "90210", "90210", "97752"),
  time1 = c(20,4,10,10,5),
  wage = c(5,7,3,12,3),
  travel_type = c("drive","drive","drive","drive","drive")
) %>%
  mutate(weight = time1*wage)

# Add reverse edges with same weights
df_reverse <- data.frame(
  from = df$to,
  to = df$from,
  time1 = df$time1,
  wage = df$wage,
  weight = df$weight,
  travel_type = df$travel_type
)

edges <- rbind(df, df_reverse) 
edges <- edges %>%
  mutate(id = as.integer(c(1:nrow(edges))))

head(edges)
```

##Step 3: Make the GRAPH
```{r}
# Convert to graph
library(igraph)
graph <- graph_from_data_frame(d=edges, vertices = nodes, directed = TRUE)

# Set the weights, ID's
E(graph)$weight #<- df_both$weight

E(graph)$travel_type

# set the vertex.id attribute for each node
V(graph)$id <- V(graph)$name

# check the vertex.id attribute
V(graph)$id

# Plot the graph (optional)
plot(graph)
```


```{r}
# Find the shortest path between two nodes using Dijkstra's algorithm
start <- "95758"
end <- "90210"

# Extract the node IDs of the shortest path
path_nodes <- V(graph)$name[get.shortest.paths(graph, from = start, to = end)$vpath[[1]]]

path_edges <- shortest_paths(graph, from = start, to = end, output = "both")$epath[[1]]

# Print the shortest path and its length
cat("Shortest path from", start, "to", end, ":", paste(path_nodes, collapse = " -> "), "\n")
cat("Length:", sum(E(graph)[as.integer(paste(path_edges))]$weight))
```

This is just output from the examples if we need:
```{r}
as_adjacency_matrix(graph, attr="weight")
as_data_frame(graph, what="edges")
```
example of graph edge output we can get
```{r}
E(graph)$id
E(graph)$weight
E(graph)$travel_type
```
example of node output we can get
```{r}
V(graph)$id
V(graph)$node_type
```


## Make a function to take a df with start zipcodes and end zipcodes, and the graph and spit out a new column for the shortest length to get there
```{r}
find_shortest_paths <- function(df, graph) {
  # Initialize a data frame to store the results
  result <- data.frame(start = character(),
                       end = character(),
                       shortest_path = character(),
                       weight_sum = numeric(),
                       stringsAsFactors = FALSE)
  
  # Loop over each row in the input data frame
  for (i in 1:nrow(df)) {
    # Extract the start and end nodes from the current row
    start <- df[i, 1]
    end <- df[i, 2]
    
    # Find the shortest path between the start and end nodes
    path <- shortest_paths(graph, from = start, to = end, output = "both")
    path_nodes <- V(graph)$name[get.shortest.paths(graph, from = start, to = end)$vpath[[1]]]
    shortest_path <- paste(path_nodes, collapse = " -> ")
    
    # Find the sum of the weights along the shortest path
    path_edges <- shortest_paths(graph, from = start, to = end, output = "both")$epath[[1]]
    weight_sum <- sum(E(graph)[path_edges]$weight)
    
    # Add the results to the output data frame
    result <- rbind(result, data.frame(start = start, end = end, shortest_path = shortest_path, weight_sum = weight_sum))
  }
  
  return(result)
}

```

lets test this out: Create example df
```{r}
test_df <- data.frame(start = c("95758", "90210", "93405"),
                 end = c("90210", "97752", "90210"),
                 stringsAsFactors = FALSE)

find_shortest_paths(test_df,graph)
```
Wow this worked hell yeah man.

# Add in code to now adjust for wages:

This is the part where I want to be able to have the weights change based off the wage of the employee.
I want to build a function that calculates a new graph based off an employee's wage and find the shortest path from their start location to end location, returning the shortest path like above and the sum of the weights.

```{r}
# Sample data frame
df2 <- data.frame(
  from = c("95758", "95758", "93405", "97752", "95758"),
  to = c("90210", "93405", "90210", "90210", "97752"),
  fixed_cost = c(20,4,10,10,5),
  variable_cost = c(1,2,1,3,1),
  travel_type = c("drive","drive","drive","drive","drive")
  #weight = ignore1*ignore2
  #weight = c(100, 28, 30, 120, 15)
) #%>%
  #mutate(weight = time1*wage)

# Add reverse edges with same weights
df2_reverse <- data.frame(
  from = df$to,
  to = df$from,
  fixed_cost = df2$fixed_cost,
  variable_cost = df2$variable_cost,
  travel_type = df2$travel_type
)

edges2 <- rbind(df2, df2_reverse) 
edges2 <- edges2 %>%
  mutate(id = as.integer(c(1:nrow(edges2))))

head(edges2)
```
make df2
```{r}
test_df2 <- data.frame(start = c("95758", "90210", "93405"),
                 end = c("90210", "97752", "90210"),
                 wage = c(1,2,3),
                 stringsAsFactors = FALSE)
head(test_df2)
```
The above is what I want my sample input data to look like, the output of the function should be the extra two columns for the employee. nrow should be the same after the function is applied



## Function accounting for different wages of employees

__Try number 1__
```{r}
library(igraph)

calculate_shortest_path <- function(edges, nodes, df2) {
  # Initialize a data frame to store the results
  result <- data.frame(start = character(),
                       end = character(),
                       shortest_path = character(),
                       weight_sum = numeric(),
                       stringsAsFactors = FALSE)

  
  # For each row in df2
  for (i in 1:nrow(df2)) {
    # Extract start and end nodes
    start_node <- as.character(df2[i, "start"])
    end_node <- as.character(df2[i, "end"])
    
    # Extract the wage from df2
    wage <- df2[i, "wage"]
    
    # Add a new column to edges with the wage value
    edges$wage <- wage
    
    # Calculate the weight based on fixed_cost, variable_cost and wage
    edges<- edges %>%
      mutate(weight = (variable_cost * wage) + fixed_cost,
             id = as.integer(c(1:nrow(edges))))
    
    # Create an igraph from the edges dataframe
    graph <- graph_from_data_frame(d=edges, vertices = nodes, directed = TRUE)
    
    # Find the shortest path between the start and end nodes
    path <- shortest_paths(graph, from = start, to = end, output = "both")
    path_nodes <- V(graph)$name[get.shortest.paths(graph, from = start, to = end)$vpath[[1]]]
    shortest_path <- paste(path_nodes, collapse = " -> ")
    
    # Find the sum of the weights along the shortest path
    path_edges <- shortest_paths(graph, from = start, to = end, output = "both")$epath[[1]]
    weight_sum <- sum(E(graph)[path_edges]$weight)
    
    
    # Calculate shortest path and weight
    #shortest_path <- shortest_paths(graph, from = start_node, to = end_node, output = "both")
    #path_nodes <- shortest_path$vpath[[1]]$name
    #weight_sum <- sum(shortest_path$epath[[1]]$weight)
    
    # Add the result to the output dataframe
    result <- rbind(result, data.frame(start = start, end = end, wage = wage, shortest_path = path_nodes, weight_sum = weight_sum))
  }
  
  return(result)
}
```

Try 1: why is this the output; help with the code I do not understand why it is returning so many rows. I dont think my loop is doing what I want
```{r}
calculate_shortest_path(edges2,nodes,test_df2)
```

_I dont really know what is happening wrong above here to fix. Need help with my loops and/or iteration._



__Try number 2__
```{r}
calculate_shortest_path <- function(edges, nodes, df2) {
  result <- data.frame(start_node = character(nrow(df2)), end_node = character(nrow(df2)), 
                       shortest_path = I(list(rep(NA, nrow(df2)))), weight_sum = numeric(nrow(df2)), stringsAsFactors = FALSE)
  
  # For each row in df2
  for (i in 1:nrow(df2)) {
    # Extract start and end nodes
    start_node <- as.character(df2[i, "start_node"])
    end_node <- as.character(df2[i, "end_node"])
    
    # Extract the wage from df2
    wage <- df2[i, "wage"]
    
    # Add a new column to edges with the wage value
    edges$wage <- wage
    
    # Calculate the weight based on fixed_cost, variable_cost and wage
    edges<- edges %>%
      mutate(weight = (variable_cost * wage) + fixed_cost,
             id = as.integer(c(1:nrow(edges))))
    
    # Create an igraph from the edges dataframe
    graph <- graph_from_data_frame(d=edges, vertices = nodes, directed = TRUE)
    
    # Find the shortest path between the start and end nodes
    path <- shortest_paths(graph, from = start, to = end, output = "both")
    path_nodes <- V(graph)$name[get.shortest.paths(graph, from = start, to = end)$vpath[[1]]]
    shortest_path <- paste(path_nodes, collapse = " -> ")
    
    # Find the sum of the weights along the shortest path
    path_edges <- shortest_paths(graph, from = start, to = end, output = "both")$epath[[1]]
    weight_sum <- sum(E(graph)[path_edges]$weight)
    
    # Assign the result to the output dataframe
    result[i, c("start_node", "end_node", "shortest_path", "weight_sum")] <- list(start_node, end_node, path_nodes, weight_sum)
  }
  
  return(result)
}

```

```{r}
calculate_shortest_path(edges2,nodes,test_df2)
```
Also did not work.


Below is me trying to work through the steps of the function with examples

```{r}
edgesA <- edges2 %>%
  mutate(wage = 189,
         weight = (variable_cost * wage) + fixed_cost)
graph2 <- graph_from_data_frame(d=edgesA, vertices = nodes, directed = TRUE)
E(graph2)$weight
E(graph2)
```

```{r}
    # Calculate shortest path and weight
    shortest_path <- shortest_paths(graph2, from = "95758", to = "90210", output = "both")
    path_nodes <- shortest_path$vpath[[1]]$name
    weight_sum <- sum(shortest_path$epath[[1]]$weight)
    
path_nodes
weight_sum
```


```{r}
plot(graph2)
```










